---
title: "Install and use package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Install and use package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  eval = FALSE
)
```

## Install the `NEesp` package

Install the versioned package:
```{r}
remotes::install_github("NOAA-EDAB/NEesp", ref = "v0.1.0")
```

Or install the most recent development package:
```{r}
remotes::install_github("NOAA-EDAB/NEesp", ref = "dev")
```

## Using the `NEesp` package

### Indicator reports

#### Run a report using package data and package template
```{r, eval = FALSE}
NEesp::render_ind_report("Black sea bass")
```

#### Run a report using package data and a local template (from `inst/indicator_bookdown_template` folder)
```{r, eval = FALSE}
# first source scripts if needed

NEesp::render_ind_report("Black sea bass", 
                         input = here::here("inst/indicator_bookdown_template"),
                         params_to_use = list(
                           species_ID = "Black sea bass",
                           ricky_survey_data = NEesp::bio_survey,
                           path = here::here("action_reports/Black sea bass", "figures//"),
                           save = TRUE
                         ), 
                         trouble = TRUE)
```

### Regression reports

#### Run a report using package data and package template
```{r, eval = FALSE}
NEesp::render_reg_report(stock_var = "Black sea bass",
                         epus_var = "MAB",
                         region_var = "Mid",
                         remove_var = FALSE,
                         lag_var = 0,
                         parent_folder = "zero_lag")
```

#### Run a report using package data and a local template (from "correlation_bookdown_template-dev" folder)
```{r, eval = FALSE}
NEesp::render_reg_report(stock_var = "Black sea bass",
                         epus_var = "MAB",
                         region_var = "Mid",
                         remove_var = FALSE,
                         lag_var = 0,
                         parent_folder = "zero_lag",
                         input = here::here("inst/correlation_bookdown_template"),
                         trouble = TRUE)
```



