---
title: "Use NEesp with the NEespShiny app"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use NEesp with the NEespShiny app}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  eval = FALSE
)
```

The [`NEespShiny` package](https://github.com/NOAA-EDAB/NEespShiny) is a Graphical User Interface for the `NEesp` package. 

## Install the `NEespShiny` package

Install the versioned package:
```{r}
remotes::install_github("NOAA-EDAB/NEespShiny", ref = "v0.1.0")
```

Or install the most recent development package:
```{r}
remotes::install_github("NOAA-EDAB/NEespShiny", ref = "dev")
```

## Using the `NEespShiny` package

The `NEespShiny` package has only one front-end function, which initiates the app.
```{r}
NEespShiny::run_NEesp()
```

Within the app, the user can input different parameters to use with the report templates. Output files can be downloaded as a zipped folder, which contains a .doc of the report, an images folder (.png), and a data folder (.csv/.RDS).

## Troubleshooting
- Override package installation errors from warnings: 
```{r}
Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS="true")
```

- You must also have the [`NEesp` package](https://github.com/NOAA-EDAB/NEesp) installed to run the `NEespShiny` app. Explicitly install `NEesp`:
```{r}
remotes::install_github("NOAA-EDAB/NEesp", ref = "v0.1.0")
```