---
title: "Using EDAB Utilities Functions for NE ESPs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using EDAB Utilities Functions for NE ESPs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  eval = FALSE
)
```


## EDAB Utilities 

This vignette demonstrates use of functions within the EDABUtilities package (https://github.com/NEFSC/READ_EDAB_Utilities/tree/main) in examining ecosystem data for ESPs. The example functions here apply input data (as .nc files) and shape files utilized in the golden tilefish ESP. 

### Crop netCDFs of bottom temperature and bottom salinity to shape file of BTS strata

Crops a 2D netCDF based on the extent of a shape file.

```{r}
crop_bottomT <- EDABUtilities::crop_nc_2d(data.in = c(here::here('data','glorys_bottomT.nc')),
                            output.files = c(here::here('data','crop_glorys_bottomT.nc')),
                            shp.file = here::here('shapefiles','BTS_STRATA.shp'),
                            #area.names = c('01030'),
                            var.name = 'bottomT',
                            write.out = F)

terra::plot(crop_bottomT[[1]])
```

```{r}
crop_bottomS <- EDABUtilities::crop_nc_2d(data.in = c(here::here('data','glorys_bottomS.nc')),
                            output.files = c(here::here('data','crop_glorys_bottomS.nc')),
                            shp.file = here::here('shapefiles','BTS_STRATA.shp'),
                            var.name = 'bottomS',
                            write.out = F)

terra::plot(crop_bottomS[[1]])

```


### Make 2d summary time series: mean monthly bottom temperature in golden tilefish BTS strata 

Provides summary statistics of 2d gridded data as time series by area.

```{r}
monthly_bt <- EDABUtilities::make_2d_summary_ts(data.in = c(here::here('data','glorys_bottomT.nc')),
                            output.files = c(here::here('data','GLORYS_monthly_bottomT.nc')),
                            shp.file = here::here('shapefiles','BTS_STRATA.shp'),
                            var.name = 'bottomT',
                            area.names = c('01030', '01040', '01070', '01080', '01110', '01120', 
                                           '01140', '01150', '01670', '01680', '01710', '01720', 
                                           '01750', '01760'),
                            statistic = 'mean',
                            agg.time = 'days',
                            tz = NA,
                            touches = TRUE,
                            write.out = F)

ggplot2::ggplot(monthly_bt[[1]],ggplot2::aes(x = time, y= value))+
  ggplot2::geom_line(lwd = 1)+ 
  ggplot2::labs(title = 'Monthly Bottom Temperature in GTF Strata') +
  ggplot2::xlab('Month') +
  ggplot2::ylab('Bottom Temperature') +
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~area)
```

### Make 2d summary time series: mean daily bottom salinity in golden tilefish BTS strata 

Provides summary statistics of 2d gridded data as time series by area.

```{r}
daily_bS <- EDABUtilities::make_2d_summary_ts(data.in = c(here::here('data','glorys_bottomS.nc')),
                   output.files = c(here::here('data','GLORYS_daily_bottomS.nc')),
                   shp.file = here::here('shapefiles','BTS_STRATA.shp'),
                   var.name = 'bottomS',
                   area.names = c('01030', '01040', '01070', '01080', '01110', '01120', 
                                           '01140', '01150', '01670', '01680', '01710', '01720', 
                                           '01750', '01760'),
                   statistic = 'mean',
                   agg.time = 'days',
                   tz = NA,
                   touches = TRUE,
                   write.out = F)

ggplot2::ggplot(daily_bS[[1]],ggplot2::aes(x = time, y= value))+
  ggplot2::geom_line(lwd = 1)+ 
  ggplot2::labs(title = 'Daily Bottom Salinity in GTF Strata') +
  ggplot2::xlab('Days') +
  ggplot2::ylab('Bottom Salinity') +
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~area)
```

### Make 2d gridded time series: calculate number of days (nd) above a bottom temperature threshold (14C) in the golden tilefish BTS strata

Provides a data frame of degree-day family statistics

```{r}
dd <- EDABUtilities::make_2d_deg_day_ts(data.in = c(here::here('data','glorys_bottomT.nc')),
                              output.files = c(here::here('data','GLORYS_monthly_bottomT.nc')),
                              shp.file = here::here('shapefiles','BTS_STRATA.shp'),
                              var.name = 'bottomT',
                              area.names = c('01030', '01040', '01070', '01080', '01110', '01120', 
                                            '01140', '01150', '01670', '01680', '01710', '01720', 
                                            '01750', '01760'),
                   ref.value = 14,
                   type = 'above',
                   statistic = 'nd',
                   write.out = F)

dd_df = dplyr::bind_rows(dd)

ggplot2::ggplot(dd_df,ggplot2::aes(x = ls.id, y = value, fill = area))+
  ggplot2::geom_bar(stat = 'identity', position = ggplot2::position_dodge(width = 1.0))+
  ggplot2::labs(title = 'Number of days above 14C', fill = "Strata Number") +
  ggplot2::xlab('GTF Strata') +
  ggplot2::ylab('Days') +
  ggplot2::theme_bw()
```

