## Trophic indicators

### Chlorophyll

#### Annual chlorophyll
```{r, chl-annual}
test <- ecodata::chl_pp %>%
  dplyr::mutate(Var = Var %>%
    stringr::str_replace_all("_", "\n") %>%
    stringr::str_replace_all(" ", "\n")) %>%
  dplyr::filter(
    stringr::str_detect(Var, "ANNUAL"),
    stringr::str_detect(Var, "CHLOR")
  ) %>%
  dplyr::mutate(Time = Time %>%
    stringr::str_replace("A_", "") %>%
    as.numeric())
NEesp::render_indicator(test, lab = names[1])

# data_prep(stock, test, lag = params$lag) %>% knitr::kable()
```

#### Monthly chlorophyll
```{r, chl-month}
test <- ecodata::chl_pp %>%
  dplyr::mutate(Var = Var %>%
    stringr::str_replace_all("_", "\n") %>%
    stringr::str_replace_all(" ", "\n")) %>%
  dplyr::filter(
    stringr::str_detect(Var, "MONTHLY"),
    stringr::str_detect(Var, "CHLOR")
  ) %>%
  dplyr::mutate(
    Time = Time %>%
      stringr::str_replace("M_", ""),
    Year = Time %>%
      stringr::str_trunc(width = 4, ellipsis = ""),
    Month = Time %>%
      stringr::str_trunc(width = 2, side = "left", ellipsis = ""),
    Var = paste(Var, "month", Month, sep = "\n")
  ) %>%
  dplyr::select(-Time) %>%
  dplyr::rename(Time = Year)
NEesp::render_indicator(test, lab = names[1])
```

#### Weekly chlorophyll
```{r, chl-week}
test <- ecodata::chl_pp %>%
  dplyr::mutate(Var = Var %>%
    stringr::str_replace_all("_", "\n") %>%
    stringr::str_replace_all(" ", "\n")) %>%
  dplyr::filter(
    stringr::str_detect(Var, "WEEKLY"),
    stringr::str_detect(Var, "CHLOR")
  ) %>%
  dplyr::mutate(
    Time = Time %>%
      stringr::str_replace("W_", ""),
    Year = Time %>%
      stringr::str_trunc(width = 4, ellipsis = ""),
    Week = Time %>%
      stringr::str_trunc(width = 2, side = "left", ellipsis = ""),
    Var = paste(Var, "week", Week, sep = "\n")
  ) %>%
  dplyr::select(-Time) %>%
  dplyr::rename(Time = Year)
NEesp::render_indicator(test, lab = names[1])
```

### Primary production

#### Annual primary production
```{r, pp-annual}
test <- ecodata::chl_pp %>%
  dplyr::mutate(Var = Var %>%
    stringr::str_replace_all("_", "\n") %>%
    stringr::str_replace_all(" ", "\n")) %>%
  dplyr::filter(
    stringr::str_detect(Var, "ANNUAL"),
    stringr::str_detect(Var, "PPD")
  ) %>%
  dplyr::mutate(Time = Time %>%
    stringr::str_replace("A_", "") %>%
    as.numeric())
NEesp::render_indicator(test, lab = names[1])

# data_prep(stock, test, lag = params$lag) %>% knitr::kable()
```

#### Monthly primary production
```{r, pp-month}
test <- ecodata::chl_pp %>%
  dplyr::mutate(Var = Var %>%
    stringr::str_replace_all("_", "\n") %>%
    stringr::str_replace_all(" ", "\n")) %>%
  dplyr::filter(
    stringr::str_detect(Var, "MONTHLY"),
    stringr::str_detect(Var, "PPD")
  ) %>%
  dplyr::mutate(
    Time = Time %>%
      stringr::str_replace("M_", ""),
    Year = Time %>%
      stringr::str_trunc(width = 4, ellipsis = ""),
    Month = Time %>%
      stringr::str_trunc(width = 2, side = "left", ellipsis = ""),
    Var = paste(Var, "month", Month, sep = "\n")
  ) %>%
  dplyr::select(-Time) %>%
  dplyr::rename(Time = Year)
NEesp::render_indicator(test, lab = names[1])
```

#### Weekly primary production
```{r, pp-week}
test <- ecodata::chl_pp %>%
  dplyr::mutate(Var = Var %>%
    stringr::str_replace_all("_", "\n") %>%
    stringr::str_replace_all(" ", "\n")) %>%
  dplyr::filter(
    stringr::str_detect(Var, "WEEKLY"),
    stringr::str_detect(Var, "PPD")
  ) %>%
  dplyr::mutate(
    Time = Time %>%
      stringr::str_replace("W_", ""),
    Year = Time %>%
      stringr::str_trunc(width = 4, ellipsis = ""),
    Week = Time %>%
      stringr::str_trunc(width = 2, side = "left", ellipsis = ""),
    Var = paste(Var, "week", Week, sep = "\n")
  ) %>%
  dplyr::select(-Time) %>%
  dplyr::rename(Time = Year)
NEesp::render_indicator(test, lab = names[1])
```

### Zooplankton

#### Spring zooplankton abundance by species
```{r, zoo-oi-spring}
test <- ecodata::zoo_oi %>%
  dplyr::filter(
    Var %>% stringr::str_detect("SD") == FALSE,
    Var %>% stringr::str_detect("spring")
  )
NEesp::render_indicator(test, lab = names[1])
```

#### Fall zooplankton abundance by species
```{r, zoo-oi-fall}
test <- ecodata::zoo_oi %>%
  dplyr::filter(
    Var %>% stringr::str_detect("SD") == FALSE,
    Var %>% stringr::str_detect("fall")
  )
NEesp::render_indicator(test, lab = names[1])
```

#### Zooplankton abundance by group
```{r, zoo-strat-abun}
test <- ecodata::zoo_strat_abun
NEesp::render_indicator(test, lab = names[1])
```

#### Abundance of Calanus CV and adults
```{r, calanus}
test <- ecodata::calanus_stage %>%
  dplyr::rename(Time = Year) %>%
  dplyr::filter(
    Var == "adt" | Var == "c5",
    Units == "No. per 100m^-3"
  ) %>%
  dplyr::group_by(Time, EPU, season) %>%
  dplyr::mutate(Val = sum(Value)) %>%
  dplyr::select(Time, Val, EPU, season) %>%
  dplyr::distinct() %>%
  dplyr::mutate(Var = paste("Calanus CV and adult", season))
NEesp::render_indicator(test, lab = names[1])
```

#### Zooplankton abundance anomaly
```{r, zoo-abund}
test <- ecodata::zoo_abund
NEesp::render_indicator(test, lab = names[1])
```

#### Zooplankton diversity index
```{r, zoo-diversity}
test <- ecodata::zoo_diversity
NEesp::render_indicator(test, lab = names[1])
```

#### Small/large copepod anomaly
```{r, zoo-sli-anom}
test <- ecodata::zoo_sli_anom
NEesp::render_indicator(test, lab = names[1])
```

#### Ichthyoplankton diversity
```{r, ich}
test <- ecodata::ichthyo_diversity
NEesp::render_indicator(test, lab = names[1])
```

### Forage fish abundance
```{r, forage}
test <- ecodata::forage_anomaly
NEesp::render_indicator(test, lab = names[1])
```
