---
title: "Black Sea Bass Bottom Temp. Update using EDABUtilities Package"
author: "Stephanie Owen"
date: "2024-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
`%>%` <- magrittr::`%>%`
```
*****Previous code calculates monthly mean and winter mean for entire area*****

## Data

### Get data through 2019 from ERDDAP
Pull the BT data from here: http://nefsctest.nmfs.local:8080/erddap/griddap/duPontavice_bottom_temp_local.html

### Get 2020-present data from GLORYS
Download here: https://data.marine.copernicus.eu/product/GLOBAL_MULTIYEAR_PHY_001_030/services 

### Calculate monthly mean
 @param data.in character vector. Bottom temperature input files (must be in .nc format)
 @param output.files character vector. Output file names corresponding to each input file.
 @param shp.file  string. Shape file = 'bsb_shape.shp'
 @param var.name string. Variable name you wish to extract = 'sea_water_temperature_at_sea_floor'
 @param area.names character vector. Names of shape file areas you want to summarize. 
 @param statistic string. Which statistic to calculate = 'mean'
 @param agg.time character. Time scale to calculate over = 'months'
```{r}
monthly_mean_bottomT <- function(...) { 
  make_2d_summary_monthly_mean <- EDABUtilities::make_2d_summary_ts(...) 
  df_monthly_mean <- make_2d_summary_monthly_mean %>%
    terra::as.data.frame(na.rm = FALSE) 
  monthly_mean_bottomT <- df_monthly_mean %>%
    tidyr::drop_na() %>%
  dplyr::mutate(dplyr::across(2:4, as.numeric),
                dplyr::across(1, lubridate::as_date),
                year = lubridate::year(time),
                month = lubridate::month(time),
                Units = c ('degC')) %>%
  dplyr::rename(value = bottomT) 
  
  return(monthly_mean_bottomT)
}
```

### Calculate winter mean
 @param data.in character vector. Bottom temperature input files (must be in .nc format)
 @param output.files character vector. Output file names corresponding to each input file.
 @param shp.file  string. Shape file = 'bsb_shape.shp'
 @param var.name string. Variable name you wish to extract = 'sea_water_temperature_at_sea_floor'
 @param area.names character vector. Names of shape file areas you want to summarize. 
 @param statistic string. Which statistic to calculate = 'mean'
 @param agg.time character. Time scale to calculate over = 'season'
```{r}
winter_mean_bottomT <- function(...) { 
  make_2d_summary_winter <- EDABUtilities::make_2d_summary_ts(...) 
  df_winter <- make_2d_summary_winter %>%
    terra::as.data.frame(na.rm = FALSE) 
  winter_bottomT <- df_winter %>%
    tidyr::drop_na() %>%
  dplyr::mutate(dplyr::across(2:4, as.numeric),
                dplyr::across(1, lubridate::as_date),
                year = lubridate::year(time),
                month = lubridate::month(time),
                Units = c ('degC')) %>%
  dplyr::filter(month %in% 1:3) %>%
  dplyr::rename(value = bottomT)  
  
  return(winter_bottomT)
}
```
 
 